Staged Abstract Interpreters

### 2019-03-24 paper/20_prelim.tex
misc

The abstract syntax we present in fact can be seen as a deep embedding of the
language -- we use datatypes to represent programs. This design choice makes us
easily add different interpretations over the AST; with the inheritance and
overriding mechanism in Scala, we may also add new language construct and reuse
the existing interpretation.

@@

\paragraph{Basic type definitions} We start with some basic type definitions.
The identifiers in the program are represented by strings. The two default
components in the interpreter are environments @Env@ and stores @Store@, i.e.,
mappings from identifiers to addressess and mappings from addresses to values,
respectively. But the @Addr@ and @Value@ are just declared as abstracrt types.

@@

\paragraph{Moandic operations} We define the return type of the interpreter as
@Ans@, which is a monad type @AnsM[_]@ wrapping the type @Value@. 
As mentioned in Section ~\ref{monadscala}, to use the @for@-comprehension
syntax, certain constraints have to be posed onto the type @AnsM@. Here, we use a
structural type @MonadOps@ to require @AnsM@ to at least implement @map@ and
@flatMap@. It is worth noting that @MonadOps@ takes another type parameter
@R[_]@ as binding-time; accordingly the outer @R@ defined in the trait is passed
to @MonadOps@. Inside of @MonadOps@, @R[_]@ wraps the data types @A@ and @B@ that are
encapsulated by the monad, but not the monad type @M@ itself. When acting as
compilers, we will also replace the monads to the ones that work on staged
values.

@@

Other primitive operations provide basic functionality for the interpreter.
The method @num@ and @close@ deal with primitive values, which lift literal
terms (e.g., lambdas) to our value representation (e.g., closures).
The method @get@ simply retrieves the value mapped from the identifier @x@ in
the environment and store. Conditionals and arithmetics is handled by @br0@
and @arith@, respectively. The methods @ap_clo@ takes a function value and an
arugment value and then does the application.

>>>>

The abstract syntax we present in fact can be seen as a deep embedding of the
language -- we use data-types to represent programs. This design choice makes us
easily add different interpretations over the AST; with the inheritance and
overriding mechanism in Scala, we may also add new language construct and reuse
the existing interpretation.

@@

\paragraph{Basic type definitions} We start with some basic type definitions.
The identifiers in the program are represented by strings. The two default
components in the interpreter are environments @Env@ and stores @Store@, i.e.,
mappings from identifiers to addresses and mappings from addresses to values,
respectively. But the @Addr@ and @Value@ are just declared as abstract types.

@@

\paragraph{Monadic operations} We define the return type of the interpreter as
@Ans@, which is an arbitrary monad type @AnsM[_]@ wrapping the type @Value@. 
As mentioned in Section ~\ref{monadscala}, to use the @for@-comprehension
syntax, certain constraints have to be posed onto the type @AnsM@. Here, we use a
structural type @MonadOps@ to require @AnsM@ to at least implement @map@ and
@flatMap@. It is worth noting that @MonadOps@ takes another type parameter
@R[_]@ as binding-time; accordingly the outer @R@ defined in the trait is passed
to @MonadOps@. Inside of @MonadOps@, @R[_]@ wraps the data types @A@ and @B@ that are
encapsulated by the monad, but not the monad type @M@ itself. When acting as
compilers, we will also replace the monads to the ones that work on staged
values.

@@

Other primitive operations provide basic functionality for the interpreter.
The method @num@ and @close@ deal with primitive values, which lift literal
terms (e.g., lambdas) to our value representation (e.g., closures).
The method @get@ simply retrieves the value mapped from the identifier @x@ in
the environment and store. Conditionals and arithmetics is handled by @br0@
and @arith@, respectively. The methods @ap_clo@ takes a function value and an
argument value and then does the application.
